Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-03-29 ThesisJP/Background/background.tex
Started on comments from Ines.

In this chapter the background information needed in order to understand the topics of this thesis and its papers is given. First, in Section \ref{sec:gpgpu}, the history and principle of general-purpose computing on graphics processing units (\nom{GPGPU}{General-purpose computing on graphics processing units}) is reviewed, followed by an introduction to medical ultrasound imaging and adaptive beamforming in Section \ref{sec:ultrasound}. Finally, the concepts of volume rendering and ultrasound field simulations are presented in Section \ref{sec:volren} and \ref{sec:field}.

\section{General-purpose computing on graphics processing units}\label{sec:gpgpu}
In a recent essay, Herb Sutter, a leading authority on software development, reviews the current state of the software and computer industry \cite{HerbSutter}. The essay bears the name "Welcome to the Jungle" and is a sequel to the essay "The Free Lunch is Over" from 2004 \cite{HerbSuttera}. The two essays spins around the fact that manufactures of central processing units (CPUs) hit a frequency-wall in the beginning of the 21st century. Until then, all mainstream computer programs where typically running in a single thread on a single CPU \textit{processing core}, and a programmer could expect the software to annually gain performance without touching the code. This was the "free-lunch"-era as depicted in Fig.\,\ref{fig:jungle}. The increase in processing power came mostly from an ever-growing clock frequency. However, power consumption and heat generation where also growing at the same rate, until the level of cooling required finally became to much in 2004.  From that point, chip manufactures had to pursue a different direction  than just increasing the clock frequency. The immediate solution was to embed several cores in one CPU. This insured continuation of Moore's law, since the number of transistors per chip could continue to grow. For programmers, multi-core CPUs meant that computer programs now had to be multi-threaded in order to annually gain performance. It was a concurrency revolution, as Herb Sutter named it, and  the "free lunch" was over.

>>>>

This chapter gives the background information needed in order to understand the topics of this thesis and its papers. First, in Section \ref{sec:gpgpu}, the history and the principle of general-purpose computing on graphics processing units (\nom{GPGPU}{General-purpose computing on graphics processing units}) are reviewed, followed by an introduction to medical ultrasound imaging and adaptive beamforming in Section \ref{sec:ultrasound}. Finally, the concepts of volume rendering and ultrasound field simulations are presented in Section \ref{sec:volren} and \ref{sec:field}.

\section{General-purpose computing on graphics processing units}\label{sec:gpgpu}
In a recent essay, Herb Sutter, a leading authority on software development, reviews the current state of the software and computer industry \cite{HerbSutter}. The essay bears the name "Welcome to the Jungle" and is a sequel to the essay "The Free Lunch is Over" from 2004 \cite{HerbSuttera}. The two essays spin around the fact that manufactures of central processing units (CPUs) hit a frequency-wall in the beginning of the 21st century. Until then, all mainstream computer programs where typically running in a single thread on a single CPU \textit{processing core}, and a programmer could expect the software to annually gain performance without touching the code. This was the "free-lunch"-era as depicted in Fig.\,\ref{fig:jungle}. The increase in processing power came mostly from an ever-growing clock frequency. However, power consumption and heat generation where also growing at the same rate, until the level of cooling required finally became to much in 2004.  From that point, chip manufactures had to pursue a different direction  than just increasing the clock frequency. The immediate solution was to embed several cores in one CPU. This insured continuation of Moore's law, since the number of transistors per chip could continue to grow. For programmers, multi-core CPUs meant that computer programs now had to be multi-threaded in order to annually gain performance. It was a concurrency revolution, as Herb Sutter named it, and  the "free lunch" was over.
